Beyond the Iceberg Hypothesis: \\Opening the Black Box of Transport Costs

### 2016-09-23 redaction/Presentations/SeminaireLEDa_Sept2016/slides_LEDa.tex
changements

%\usepackage[dark]{beamerthemesidebar}
%\usepackage{beamerthemeshadow}
%\usepackage{beamerthemesplit}
%\usepackage{beamerthemebars}
%\usepackage{beamerthemeclassic}
\newcommand{\spac}{\rule[0mm]{0mm}{2mm}}
\usepackage{pgf,pgfarrows,pgfnodes,pgfautomata,pgfheaps,pgfshade}
\usepackage{amsmath,amssymb}
\usepackage[english]{babel}
\usepackage{epic,eepic,bez123}
\usepackage{ bbold }
\usepackage{ upgreek }
%\usepackage[T1]{fontenc}
%\usepackage[ansinew]{inputenc}
%\usepackage[latin1]{inputenc}
%\usepackage{babel,varioref}

@@

\begin{frame}
\frametitle{Motivation (cont')}
\begin{itemize}
\item Standard modeling of trade costs: As an ad-valorem tax-equivalent \vspace{0.1cm}
\begin{itemize}
\item[-] As a constant percentage of the producer price per unit traded \vspace{0.1cm}
\item[$\Leftrightarrow$] The ``iceberg cost'' hypothesis (Samuelson, 1954) \vspace{0.1cm}
\end{itemize}
\item Yet... A debated question \vspace{0.1cm}
\item Would not trade costs rather exhibit an additive structure ?  \vspace{0.1cm}
\item The structure (additive vs iceberg) of transport costs, not anecdotal \vspace{0.1cm}
\begin{itemize}
\footnotesize{
\item[-] Additive trade costs \& the pattern of trade flows (Alchian \& Allen, 1964) \vspace{0.1cm}
\item[-] Strong normative implications (Sorensen, 2014)\vspace{0.1cm}
\item[-] The additive structure of trade costs is supported by recent empirical evidence (Irrarazabal et al. 2014, Hummels \& Skiba, 2004)  \vspace{0.1cm}}
\end{itemize}
\item[$\Rightarrow$] Trade costs are likely to display an additive component, but precisely... by how much? \vspace{0.1cm}
\end{itemize}
\textbf{One objective of the paper: Provide an answer to this question}
\end{frame}

\begin{frame}
\frametitle{Our paper in 3 questions (and 3 answers)}
An empirical decomposition of the structure of transport costs \vspace{0.1cm}
\begin{itemize}
\item[(1)] What is the size of the iceberg and the additive transport costs? \vspace{0.1cm}
\item[$\Rightarrow$] Provide a quantitative measure of both, using US imports data \vspace{0.1cm}
\begin{itemize}
\footnotesize{
\item[-] Iceberg cost: 2.5\% of the export price in air, and 3.2\% in vessel (mean value over 1974-2013) \vspace{0.1cm}
\item[-] Additive cost: 1.8\% and 2.9\% of the export price, in air / vessel  \vspace{0.1cm}
}
\end{itemize}
\item[(2)] What do we lose by skipping the additive part of transport costs? \vspace{0.1cm}
\item[$\Rightarrow$] We lose much: With the additive term included, \vspace{0.1cm}
\begin{itemize}
\footnotesize{
\item[-] The estimated iceberg component is reduced by a factor of 2 \vspace{0.1cm}
\item[-] A significantly better ``goodness-of-fit'' \vspace{0.1cm}
}
\end{itemize}
\item[(3)] How have international transport costs evolved over time? \vspace{0.1cm}
\item[$\Rightarrow$] The importance of excluding composition effects \vspace{0.1cm}
\begin{itemize}
\footnotesize{
\item[-]\textit{Pure} transport costs have fallen since 1985 (not 1974), by $\simeq$ 40\% \vspace{0.1cm}
}
\end{itemize}
\item[$\Rightarrow$] The importance of including the additive component \vspace{0.1cm}
\begin{itemize}
\footnotesize{
\item[-] Not much difference between air and sea transport, $\neq$ Hummels (2007) and Behar \& Venables (2010) \vspace{0.1cm}
%\item[-] Marked difference in the evolution of additive and multiplicative costs in Air
%\end{itemize}
}
\end{itemize}
\end{itemize}
\end{frame}

@@

\begin{frame}
\frametitle{Empirical specification (1)}
\textbf{The estimated equation}
\begin{itemize}
\item Relate the import price $p$ to the export price $\widetilde{p}$ given both additive (per-kg) costs $t$ and ad-valorem costs $\tau$:
$$p = \tau \widetilde{p} + t, \qquad \text{with}\quad \tau \geq 1,\quad t \geq 0$$
\item For product $k$, from country $i$  \vspace{0.1cm}
\item Rewrite to get:
$$\frac{p_{ik}}{\widetilde{p}_{ik}} -1 = \tau_{ik} -1 +\frac{t_{ik}}{ \widetilde{p}_{ik}}$$
\item[$\Rightarrow$] Estimate this equation for each year over 1974-2013 (at the $k=$ 3-digit classification level) \vspace{0.1cm}
\begin{itemize}
\item[-] The equation is also year- and mode (air or vessel)- specific
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Empirical specification (2)}
\textbf{The estimation strategy}
\begin{itemize}
\item Two assumptions (as in Irrarazabal et al., 2015) \vspace{0.1cm}
\begin{itemize}
\item[-] Both iceberg and additive costs are separable between the origin country $i$ and the product $k$ dimensions \vspace{0.1cm}
\item[-] Separability in a multiplicative manner for ad-valorem costs and additive manner for per-kg costs \vspace{0.1cm}
\item[$\Leftrightarrow$] Write $t_{ik}$ and $\tau_{ik}$ as:
\begin{equation}
 \tau_{ik} = \tau_{i} \times \tau_{k}, \qquad t_{ik} = t_{i} + t_{k} \label{eq:specifTC}
 \end{equation}
\end{itemize}
\item Given the constraint $\frac{p_{ik}}{\widetilde{p}_{ik}} -1$, the error term should be always positive and multiplicative \vspace{0.1cm}
\item[$\Rightarrow$] The estimated equation becomes:
\footnotesize
$$\frac{p_{ik}}{\widetilde{p}_{ik}}-1 =\left(\tau_{i} \times \tau_{k} -1+\frac{t_{i} + t_{k}}{\widetilde{p}_{ik}} \right)\times \exp(\epsilon_{ik})$$
\normalsize
\begin{itemize}
\item[-] With $\epsilon_{ik}$ following a normal law centered on 0.
\end{itemize}

@@

\begin{frame} [label=slide_method]
\begin{itemize}
\item Taking logs, we finally estimate the following equation
\footnotesize
\begin{equation}
\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i} \times \tau_{k}+\frac{t_{i} + t_{k}}{\widetilde{p}_{ik}}-1 \right) + \epsilon_{ik} \label{eq:est_equation}
\end{equation}
\normalsize
\item A non-linear equation (due to the additive costs)  \vspace{0.1cm}
\item[$\Rightarrow$] Estimation using non-linear squares \hyperlink{app_method}{\beamergotobutton{More}} \vspace{0.2cm}
\item How to characterize the importance of additive costs relatively to iceberg? \vspace{0.1cm}
\item[$\Rightarrow$] Estimate Equation (\ref{eq:est_equation}) constraining $t=0$ \vspace{0.1cm}
\footnotesize
\begin{equation}
\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i} \times \tau_{k}-1 \right) + \epsilon^{ice}_{ik} \label{eq:model_nlI}
\end{equation}
\normalsize
\item Estimate \textit{two} equations, for each year over 1974-2013, by transport mode (air/ocean)\vspace{0.1cm}
\begin{itemize}
\item[-] With additive costs included: Equation (\ref{eq:est_equation})  \vspace{0.1cm}
\begin{itemize}
\item[$\Rightarrow$] By year and transport mode, $\simeq$ 800 fixed effects to estimate   \vspace{0.1cm}
\end{itemize}
\item[-] With only ad-valorem costs: Equation (\ref{eq:model_nlI})
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame} 
\begin{itemize}
\item After running the estimates, we re-built: \vspace{0.1cm}
\begin{itemize}
\item[-] With additive costs:
$$\widehat{\tau}^{adv}_{ik} = \widehat{\tau_{i}} \times \widehat{\tau_{k}}, \qquad \widehat{t}^{add}_{ik} = \widehat{t}_{i} + \widehat{t}_{k}$$
\item[-] With only iceberg costs:
$$\widehat{\tau}^{ice}_{ik} = \widehat{\tau_{i}} \times \widehat{\tau_{k}}$$
\end{itemize}
\item  Taking the average over the product-country dimension, we finally get (by year and transport mode): \vspace{0.1cm}
\begin{itemize}
\item[-] When additive costs are included: $\widehat{\tau}^{adv}$, $\widehat{t}^{add}$  \vspace{0.1cm}
\item[-] With only iceberg costs: $\widehat{\tau}^{ice}$
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}[label=slide_results_summary]
\frametitle{Result 1: Estimating transport costs over time}
For both the ad-valorem and the additive components \vspace{0.1cm}
\begin{itemize}
\item[-] \footnotesize{Average values over 1974-2013, in percent of the export price \hyperlink{app_results_summary}{\beamergotobutton{More}}  } \vspace{0.1cm}
\end{itemize}
\begin{table}[htbp]
  \centering
  \scriptsize{
  %\caption{Transport costs estimates: Summary \label{tab:summary_results}}
  \begin{center}
    \begin{tabular}{l|cc|cc}
      \hline \hline
    %\multicolumn{5}{c}{Mean value over 1974-2013}   \\
    \# digit & \multicolumn{2}{c}{3 digits} & \multicolumn{2}{c}{4 digits} \\ \hline
    Mode  & Vessel & Air & Vessel & Air \\ \hline
    \multicolumn{5}{l}{With only Ad-Valorem Costs ($\widehat{\tau}^{ice}$)}  \\ \hline
    Mean  & \bf{5.8} & \bf{5.1} & 6.0 & 4.9 \\
    Median & 5.1 & 4.2 & 5.2 & 3.7 \\ \hline
    %Std   & 0.032 & 0.042 & 0.036 & 0.045 \\
    %Min. value & 1.003 & 1.001 & 1.003 & 1.000 \\
    %Max. value & 1.304 & 1.685 & 1.408 & 2.051 \\ \hline
    \multicolumn{5}{l}{With Additive \& Ad-Valorem Costs } \\ \hline
   \multicolumn{5}{l}{\textit{Ad-valorem term} ($\widehat{\tau}^{adv}$)}\\ \hline
    Mean  & \bf{3.2} & \bf{2.5} & 3.3 & 2.4 \\
    Median & 2.8 & 1.8 & 2.8 & 1.6 \\ \hline
    %Std   & 0.023 & 0.023 & 0.025 & 0.026 \\
    %Min. value & 1.001 & 1.000 & 1.000 & 1.000 \\
    %Max. value & 1.227 & 1.474 & 1.264 & 1.537 \\ \hline
    \multicolumn{5}{l}{\textit{Additive term }($\widehat{t}^{add}/\widetilde{p}$)} \\ \hline
    Mean  & \bf{2.9} & \bf{1.8} & 2.8 & 1.9 \\
    Median & 1.9 & 0.7 & 1.7 & 0.8 \\ \hline 
    %Std   & 0.041 & 0.034 & 0.039 & 0.034 \\
    %Min. value & 0.000 & 0.000 & 0.000 & 0.000 \\
    %Max. value & 2.941 & 13.303 & 3.197 & 11.440 \\ \hline
       \multicolumn{5}{l}{\textit{Data} }\\ \hline
    Mean &5.6 &4.0 &5.7 & 4.0\\
    Median  &4.4 & 1.9 &4.5 & 2.0 \\  \hline 
    \# obs. & 29279 & 28207 & 29317 & 27680 \\ 
    \# origin country & 188 & 191 & 188 & 189 \\
    \# products & 230 & 211 & 666 & 567 \\    \hline \hline
  \end{tabular}
%\parbox[l]{8cm}{\tiny{Notes: Statistics are obtained weighting each observation by its value.}}
%. The additive term is expressed in fraction of fab price. ($^\ast$): Four 4-digit estimation: 0n selected years. ($^{\ast \ast}$): 1989 omitted in 3 digit estimation for air.}}
\end{center}}
  \end{table}%
 \end{frame}

@@

\begin{frame}
\frametitle{Excluding the composition effects of transport costs changes}
\begin{itemize}
\item For the ad-valorem component, we estimate the following equation:
\scriptsize
\begin{eqnarray}
%\widehat{\tau}_{ikt}&=&\delta\times \exp\left(\sum_{i \neq \text{AFG}}\alpha_i.\mathbb{1}_i\right).\exp\left(\sum_{k\neq \text{011}}\beta_k.\mathbb{1}_k\right).\exp\left(\sum_{t \neq 1974}\gamma_t.\mathbb{1}_t\right) .\exp\left(\epsilon_{ikt}\right) \notag \\
\ln(\widehat{\tau}_{ikt})&=&\delta +\underbrace{\sum_{i \neq \text{AFG}}\alpha_i.\mathbb{1}_i }_{\simeq \ln(\tau_i)}+ \underbrace{\sum_{k\neq \text{011}}\beta_k.\mathbb{1}_k }_{+ \simeq \ln(\tau_k)}+ \underbrace{\sum_{t \neq 1974}\gamma_t.\mathbb{1}_t}_{\text{``Pure'' TC}}+\epsilon_{ikt} \label{eq:compeffects_mult}
\end{eqnarray}
\normalsize
\begin{itemize}
\item[-] With $\widehat{\tau}_{ikt} = \widehat{\tau}^{ice}_{ikt}, \widehat{\tau}^{adv}_{ikt}$ previously obtained  \vspace{0.1cm}
\end{itemize}
\item For the additive component:
\scriptsize
\begin{eqnarray}
%\widehat{t}_{ikt}&=&\left( \prod_{i \neq \text{ARG}}  \alpha_i.\mathbb{1}_i+\prod_{k}\beta_k.\mathbb{1}_k\right).\exp\left(\sum_{t \neq 1974}\gamma_t.\mathbb{1}_t\right) .\exp\left(\epsilon_{ikt}\right) \notag \\
\ln(\widehat{t}_{ijt})&=&\ln\left( \underbrace{\prod_{i \neq \text{ARG}}  \alpha_i.\mathbb{1}_i}_{\simeq \ln(t_i)}+\underbrace{\prod_{k\neq \text{011}}\beta_k.\mathbb{1}_k}_{+ \simeq t_k}\right) + \sum_{t \neq 1974}\gamma_t.\mathbb{1}_t+\epsilon_{ijt} \label{eq:compeffects_add}
\end{eqnarray}
\normalsize
\item Underlying rationale
\begin{itemize}
\item[-] Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}): Preserve our specification of the ad-valorem and the additive costs (Equation (\ref{eq:specifTC})) \vspace{0.1cm}
\item[-] Equation (\ref{eq:compeffects_mult}) estimated using OLS, Equation (\ref{eq:compeffects_add}) using non-linear least squares (by transport mode) \vspace{0.1cm}
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\begin{itemize}
\item Exclude the composition effects of transport costs changes  \vspace{0.1cm}
\item[$\Leftrightarrow$] Isolate the change in the time dimension  \vspace{0.1cm}
\begin{itemize}
\item[-] From the ad-valorem component estimation (Equation (\ref{eq:compeffects_mult})), build the variable $\Gamma_t$ ($\forall~t \geq 1974$):
\begin{equation*}
\Gamma_t = 100.\frac {\bar{\tau}_{1974}.\exp(\gamma_t)-1} {\bar{\tau}_{1974}-1}
\end{equation*}
\begin{itemize}
\item[$\ast$] with $\bar{\tau}_{1974} = ???$
\end{itemize}
\item[-] For the additive cost, we build the variable ($\forall~t \geq 1974$)
$$\Gamma^{add}_t = 100 \exp(\gamma_t)$$
\end{itemize}
\item The $\Gamma^{add}_t$ and $\Gamma_t$ series: Interpretation in percentage changes
\begin{itemize}
\item[] with an initial value of 100 for $t=1974$
\end{itemize}
\end{itemize}

@@

\begin{frame}
\frametitle{Conclusion (2)}

@@

\begin{frame}[label=app_method]
\frametitle{More on the estimation method}
\begin{itemize}
\item The non-linear least squares method \vspace{0.1cm}
\begin{itemize}
\item[-] At the basis of the method: Approximate the model by a linear one and refine the parameters by successive iterations \vspace{0.1cm}
\item[-] The criterion for convergence: That the sum of the squares of the residuals does not increase from one iteration to the next. \vspace{0.1cm}
\end{itemize}
\item Eliminate potential influence of outliers: Exclude the 5 percent of the upper and lower tails of the distribution \vspace{0.1cm}

\end{itemize}
\hyperlink{slide_method}{\beamergotobutton{Back to slide}}
\end{frame}

>>>>

%\usepackage[dark]{beamerthemesidebar}
%\usepackage{beamerthemeshadow}
%\usepackage{beamerthemesplit}
%\usepackage{beamerthemebars}
%\usepackage{beamerthemeclassic}
\newcommand{\spac}{\rule[0mm]{0mm}{2mm}}
\usepackage{pgf,pgfarrows,pgfnodes,pgfautomata,pgfheaps,pgfshade}
\usepackage{amsmath,amssymb}
\usepackage[english]{babel}
\usepackage{epic,eepic,bez123}
\usepackage{ bbold }
\usepackage{ upgreek }
\usepackage{eurosym}
%\usepackage[T1]{fontenc}
%\usepackage[ansinew]{inputenc}
%\usepackage[latin1]{inputenc}
%\usepackage{babel,varioref}

@@

\begin{frame}
\frametitle{Motivation (cont')}
\begin{itemize}
\item Standard modeling of trade costs: As an ad-valorem tax-equivalent \vspace{0.1cm}
\begin{itemize}
\item[-] As a constant percentage of the producer price per unit traded \vspace{0.1cm}
\item[$\Leftrightarrow$] The ``iceberg cost'' hypothesis (Samuelson, 1954) \vspace{0.1cm}
\begin{itemize}
\item[-] With $p$ the import price, $\widetilde{p}$ the export price, $q$ the quantity traded and $\tau >1$ the trade cost \vspace{0.1cm}
\footnotesize
$$pq = \tau \widetilde{p}q$$
\normalsize
\end{itemize}
\end{itemize}
\item Yet... A debated question \vspace{0.1cm}
\item Would not trade costs rather exhibit an additive structure ?  \vspace{0.1cm} \vspace{0.1cm}
\begin{itemize}
\item[-] Why would it be more costly to transport from Milan to Paris, a pair of Italian shoes at price \euro 300 than a pair of Italian shoes at price \euro 50? \vspace{0.1cm}
\item[-] Pricing shipping often includes an additive component \vspace{0.1cm}
\begin{itemize}
\item[$\star$] UPS: A \$125 fee charged for a 2 pound package from Oslo to NYC (Irrarazabal et al., 2014) \vspace{0.1cm}
\end{itemize}
\item[-] Some trade policy instruments as well (quota license price)
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\begin{itemize}
\item The structure (additive vs iceberg) of transport costs, not anecdotal \vspace{0.2cm}
\begin{itemize}
\item[-] Additive trade costs play an important role in shaping the pattern of trade flows (Alchian \& Allen, 1964) \vspace{0.1cm}
\item[-] Strong normative implications, notably w.r.t. the welfare gains of trade liberalization (Sorensen, 2014)\vspace{0.1cm}
\item[-] The additive structure of trade costs is supported by recent empirical evidence (Irrarazabal et al. 2014, Hummels \& Skiba, 2004)  \vspace{0.1cm}
\end{itemize}
\item[$\Rightarrow$] Trade costs are likely to display an additive component, but precisely... by how much? \vspace{0.2cm}
\end{itemize}
\textbf{One objective of the paper: Provide an answer to this question}

\begin{frame}
\frametitle{Our paper in 3 questions (and 3 answers)}
An empirical decomposition of the structure of transport costs \vspace{0.2cm}
\begin{itemize}
\item[(1)]\textbf{ What is the size of the iceberg and the additive costs?} \vspace{0.1cm}
\item[$\Rightarrow$] Provide a quantitative measure of both, using US imports data \vspace{0.1cm}
\begin{itemize}
%\footnotesize{
\item[-] Iceberg cost: 2.5\% of the export price in air, and 3.2\% in vessel (mean value over 1974-2013) \vspace{0.1cm}
\item[-] Additive cost: 1.8\% and 2.9\% of the export price, in air / vessel  \vspace{0.2cm}
%}
\end{itemize}
\item[(2)] \textbf{ What do we lose by skipping the additive part of transport costs?} \vspace{0.1cm}
\item[$\Rightarrow$] We lose much: With the additive term included, \vspace{0.1cm}
\begin{itemize}
%\footnotesize{
\item[-] The estimated iceberg component is reduced by a factor of 2 \vspace{0.1cm}
\item[-] A significantly better ``goodness-of-fit'' \vspace{0.1cm}
%}
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\begin{itemize}
\item[(3)] \textbf{ How have international transport costs evolved over time?} \vspace{0.2cm}
\item[$\Rightarrow$] The importance of excluding composition effects \vspace{0.2cm}
\begin{itemize}
%\footnotesize{
\item[-]\textit{Pure} transport costs have fallen since 1985 (not 1974), by $\simeq$ 40\% \vspace{0.2cm}
%}
\end{itemize}
\item[$\Rightarrow$] The importance of including the additive component \vspace{0.2cm}
\begin{itemize}
%\footnotesize{
\item[-] Not much difference between air and sea transport, $\neq$ Hummels (2007) and Behar \& Venables (2010) \vspace{0.2cm}
\item[-] Marked difference in the evolution of additive and multiplicative costs in Air
%\end{itemize}
%}
\end{itemize}
\end{itemize}
\end{frame}

@@

\begin{frame}
\frametitle{Empirical specification (1)}
\textbf{The equation at the root of the estimation}
\begin{itemize}
\item Relate the import price $p$ to the export price $\widetilde{p}$ given both additive (per-kg) costs $t$ and ad-valorem costs $\tau$:
$$p = \tau \widetilde{p} + t, \qquad \text{with}\quad \tau \geq 1,\quad t \geq 0$$
\item For product $k$, from country $i$  \vspace{0.1cm}
\item Rewrite to get:
$$\frac{p_{ik}}{\widetilde{p}_{ik}} -1 = \tau_{ik} -1 +\frac{t_{ik}}{ \widetilde{p}_{ik}}$$
\item[$\Rightarrow$] Estimate this equation for each year over 1974-2013 (at the $k=$ 3-digit classification level) \vspace{0.1cm}
\begin{itemize}
\item[-] The equation is also year- and mode (air or vessel)- specific
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Empirical specification (2)}
\textbf{The estimation strategy}
\begin{itemize}
\item Two assumptions (as in Irrarazabal et al., 2015) \vspace{0.1cm}
\begin{itemize}
\item[-] Both iceberg and additive costs are separable between the origin country $i$ and the product $k$ dimensions \vspace{0.1cm}
\item[-] Separability in a multiplicative manner for ad-valorem costs and additive manner for per-kg costs \vspace{0.1cm}
\item[$\Leftrightarrow$] Write $t_{ik}$ and $\tau_{ik}$ as:
\begin{equation}
 \tau_{ik} = \tau_{i} \times \tau_{k}, \qquad t_{ik} = t_{i} + t_{k} \label{eq:specifTC}
 \end{equation}
\end{itemize}
\item Given the constraint $\frac{p_{ik}}{\widetilde{p}_{ik}} -1$, the error term should be always positive and multiplicative \vspace{0.1cm}
\item[$\Rightarrow$] The equation of interest becomes:
\footnotesize
$$\frac{p_{ik}}{\widetilde{p}_{ik}}-1 =\left(\tau_{i} \times \tau_{k} -1+\frac{t_{i} + t_{k}}{\widetilde{p}_{ik}} \right)\times \exp(\epsilon_{ik})$$
\normalsize
\begin{itemize}
\item[-] With $\epsilon_{ik}$ following a normal law centered on 0.
\end{itemize}

@@

\begin{frame} 
\begin{itemize}
\item Taking logs, we finally estimate the following equation

\begin{equation}
\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i} \times \tau_{k}+\frac{t_{i} + t_{k}}{\widetilde{p}_{ik}}-1 \right) + \epsilon_{ik} \label{eq:est_equation}
\end{equation}
\vspace{0.1cm}
\item A non-linear equation (due to the additive costs)  \vspace{0.1cm}
\item[$\Rightarrow$] Estimation using non-linear squares \vspace{0.1cm}
\begin{itemize}
\item[-] At the basis of the method: Approximate the model by a linear one and refine the parameters by successive iterations \vspace{0.1cm}
\item[-] The criterion for convergence: That the sum of the squares of the residuals does not increase from one iteration to the next. \vspace{0.1cm}
\item[-] Eliminate potential influence of outliers: Exclude the 5 percent of the upper and lower tails of the distribution \vspace{0.1cm}
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Empirical specification (3)}
\textbf{How to characterize the importance of additive costs relatively to iceberg?} \vspace{0.1cm}
\begin{itemize}
\item Estimate Equation (\ref{eq:est_equation}) constraining $t=0$ \vspace{0.1cm}
\item[$\Rightarrow$] Estimate \textit{two} equations, for each year over 1974-2013, by transport mode (air/ocean)\vspace{0.1cm}
\begin{itemize}
\item[-] With additive costs included: 
%\footnotesize
\begin{equation}
\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i} \times \tau_{k}+\frac{t_{i} + t_{k}}{\widetilde{p}_{ik}}-1 \right) + \epsilon_{ik}
\end{equation}  
%\normalsize
\begin{itemize}
\item[$\Rightarrow$] By year and transport mode, $\simeq$ 800 fixed effects to estimate   \vspace{0.1cm}
\end{itemize}
\item[-] With only ad-valorem costs:
%\footnotesize
\begin{equation}
\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i} \times \tau_{k}-1 \right) + \epsilon^{ice}_{ik} \label{eq:model_nlI}
\end{equation}
%\normalsize
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\begin{itemize}
\item After running the estimates, we re-built: \vspace{0.1cm}
\begin{itemize}
\item[-] With additive costs:
$$\widehat{\tau}^{adv}_{ik} = \widehat{\tau_{i}} \times \widehat{\tau_{k}}, \qquad \widehat{t}^{add}_{ik} = \widehat{t}_{i} + \widehat{t}_{k}$$
\item[-] With only iceberg costs:
$$\widehat{\tau}^{ice}_{ik} = \widehat{\tau_{i}} \times \widehat{\tau_{k}}$$
\end{itemize}
\item  Taking the weighted average over the product-country dimension, we finally get (by year and transport mode): \vspace{0.1cm}
\begin{itemize}
\item[-] When additive costs are included: $\widehat{\tau}^{adv}$, $\widehat{t}^{add}$  \vspace{0.1cm}
\item[-] With only iceberg costs: $\widehat{\tau}^{ice}$
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}[label=slide_results_summary]
\frametitle{Result 1: Estimating transport costs over time}
For both the ad-valorem and the additive components \vspace{0.1cm}
\begin{itemize}
\item[-] \footnotesize{Average values over 1974-2013, in percent of the export price \hyperlink{app_results_summary}{\beamergotobutton{More}}  } \vspace{0.1cm}
\end{itemize}
\begin{table}[htbp]
  \centering
  \scriptsize{
  %\caption{Transport costs estimates: Summary \label{tab:summary_results}}
  \begin{center}
    \begin{tabular}{l|cc|cc}
      \hline \hline
    %\multicolumn{5}{c}{Mean value over 1974-2013}   \\
    \# digit & \multicolumn{2}{c}{3 digits} & \multicolumn{2}{c}{4 digits} \\ \hline
    Mode  & Vessel & Air & Vessel & Air \\ \hline
    \multicolumn{5}{l}{With only Ad-Valorem Costs ($\widehat{\tau}^{ice}$)}  \\ \hline
    Mean  & \bf{5.8} & \bf{5.1} & 6.0 & 4.9 \\
    Median & 5.1 & 4.2 & 5.2 & 3.7 \\ \hline
    %Std   & 0.032 & 0.042 & 0.036 & 0.045 \\
    %Min. value & 1.003 & 1.001 & 1.003 & 1.000 \\
    %Max. value & 1.304 & 1.685 & 1.408 & 2.051 \\ \hline
    \multicolumn{5}{l}{With Additive \& Ad-Valorem Costs } \\ \hline
   \multicolumn{5}{l}{\textit{Ad-valorem term} ($\widehat{\tau}^{adv}$)}\\ \hline
    Mean  & \bf{3.2} & \bf{2.5} & 3.3 & 2.4 \\
    Median & 2.8 & 1.8 & 2.8 & 1.6 \\ \hline
    %Std   & 0.023 & 0.023 & 0.025 & 0.026 \\
    %Min. value & 1.001 & 1.000 & 1.000 & 1.000 \\
    %Max. value & 1.227 & 1.474 & 1.264 & 1.537 \\ \hline
    \multicolumn{5}{l}{\textit{Additive term }($\widehat{t}^{add}/\widetilde{p}$)} \\ \hline
    Mean  & \bf{2.9} & \bf{1.8} & 2.8 & 1.9 \\
    Median & 1.9 & 0.7 & 1.7 & 0.8 \\ \hline
    %Std   & 0.041 & 0.034 & 0.039 & 0.034 \\
    %Min. value & 0.000 & 0.000 & 0.000 & 0.000 \\
    %Max. value & 2.941 & 13.303 & 3.197 & 11.440 \\ \hline
    %   \multicolumn{5}{l}{\textit{Data} }\\ \hline
    %Mean &5.6 &4.0 &5.7 & 4.0\\
    %Median  &4.4 & 1.9 &4.5 & 2.0 \\  \hline
    \# obs. & 29279 & 28207 & 29317 & 27680 \\
    \# origin country & 188 & 191 & 188 & 189 \\
    \# products & 230 & 211 & 666 & 567 \\    \hline \hline
  \end{tabular}
%\parbox[l]{8cm}{\tiny{Notes: Statistics are obtained weighting each observation by its value.}}
%. The additive term is expressed in fraction of fab price. ($^\ast$): Four 4-digit estimation: 0n selected years. ($^{\ast \ast}$): 1989 omitted in 3 digit estimation for air.}}
\end{center}}
  \end{table}%
 \end{frame}

@@

\begin{frame}
\frametitle{Excluding the composition effects}
\begin{itemize}
\item For the ad-valorem component, we estimate the following equation:
\footnotesize
\begin{eqnarray}
%\widehat{\tau}_{ikt}&=&\delta\times \exp\left(\sum_{i \neq \text{AFG}}\alpha_i.\mathbb{1}_i\right).\exp\left(\sum_{k\neq \text{011}}\beta_k.\mathbb{1}_k\right).\exp\left(\sum_{t \neq 1974}\gamma_t.\mathbb{1}_t\right) .\exp\left(\epsilon_{ikt}\right) \notag \\
\ln(\widehat{\tau}_{ikt})&=&\delta +\underbrace{\sum_{i \neq \text{AFG}}\alpha_i.\mathbb{1}_i }_{\ln(\widetilde{\tau}_i)}+ \underbrace{\sum_{k\neq \text{011}}\beta_k.\mathbb{1}_k }_{\ln(\widetilde{\tau}_k)}+ \underbrace{\sum_{t \neq 1974}\gamma_t.\mathbb{1}_t}_{\text{``Pure'' TC}}+\epsilon_{ikt} \label{eq:compeffects_mult}
\end{eqnarray}
\normalsize
\begin{itemize}
\item[-] With $\widehat{\tau}_{ikt} = \widehat{\tau}^{ice}_{ikt}, \widehat{\tau}^{adv}_{ikt}$ previously obtained  \vspace{0.1cm}
\end{itemize}
\item For the additive component:
\footnotesize
\begin{eqnarray}
%\widehat{t}_{ikt}&=&\left( \prod_{i \neq \text{ARG}}  \alpha_i.\mathbb{1}_i+\prod_{k}\beta_k.\mathbb{1}_k\right).\exp\left(\sum_{t \neq 1974}\gamma_t.\mathbb{1}_t\right) .\exp\left(\epsilon_{ikt}\right) \notag \\
\ln(\widehat{t}_{ijt})&=&\ln\left( \underbrace{\prod_{i \neq \text{ARG}}  \alpha_i.\mathbb{1}_i}_{\widetilde{t}_i}+ \underbrace{\prod_{k}\beta_k.\mathbb{1}_k}_{\widetilde{t}_k}\right) + \underbrace{\sum_{t \neq 1974}\gamma_t.\mathbb{1}_t}_{\text{``Pure'' TC}}+\epsilon_{ijt} \label{eq:compeffects_add}
\end{eqnarray}
\normalsize
\item Underlying rationale
\footnotesize
\begin{itemize}
\item[-] Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}): Preserve our specification of the ad-valorem and the additive costs (Equation (\ref{eq:specifTC})) \vspace{0.1cm}
\item[-] Equation (\ref{eq:compeffects_mult}) estimated using OLS, Equation (\ref{eq:compeffects_add}) using non-linear least squares (by transport mode)
    \normalsize
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\begin{itemize}
\item Exclude the composition effects of transport costs changes  \vspace{0.1cm}
\item[$\Leftrightarrow$] Isolate the change in the time dimension  \vspace{0.1cm}
\begin{itemize}
\item[-] From the ad-valorem component estimation (Equation (\ref{eq:compeffects_mult})), build the variable $\Gamma_t$ ($\forall~t > 1974$):
\begin{eqnarray*}
&&\Gamma_t = 100\frac {\bar{\tau}_{t}-1} {\bar{\tau}_{1974}-1} \\
\Leftrightarrow && \Gamma_t= \frac {\bar{\tau}_{1974}.\exp(\gamma_t)-1} {\bar{\tau}_{1974}-1}
\end{eqnarray*}
\begin{itemize}
\item[$\ast$] with $\bar{\tau}_{1974} = \exp(\delta+\sum_i \alpha_i + \sum_k \beta_k$) the mean TC in 1974
\end{itemize}
\item[-] For the additive cost, we build the variable ($\forall~t > 1974$)
$$\Gamma^{add}_t = 100 \exp(\gamma_t)$$
\end{itemize}
\item The $\Gamma^{add}_t$ and $\Gamma_t$ series: Interpretation in percentage changes
\begin{itemize}
\item[] with an initial value of 100 for $t=1974$
\end{itemize}
\end{itemize}

@@

\begin{frame}
\frametitle{Conclusion : What to do next?}

@@
